---
title: "Reproduced analysis of FARS data"
author: "Tulsi Ram Gompo"
date: "11/2/2016"
output: pdf_document
---

```{r setup, include=FALSE }
knitr::opts_chunk$set(echo = FALSE)
```

**Results** 

**Percentages of drivers testing positive by drug type, sex, and year group**

```{r echo=FALSE,message=FALSE}
load("/Users/Tulsigompo/Desktop/R_project_new/data/clean_fars.RData")
source("/Users/Tulsigompo/Desktop/R_project_new/R/fars_functions.R")
library(dplyr)
library(tidyr)
library(ggthemes)
clean_fars %>%
mutate(year_cat = cut(year, breaks = c(1999, 2002, 2006, 2010),
labels = c("1999-2002", "2003-2006",
"2007-2010"),
include.lowest = TRUE, right = TRUE)) %>%
filter(!is.na(sex)) %>%
group_by(drug_type, sex, year_cat) %>%
dplyr::filter(!is.na (agecat))  %>%
summarize(n_non_missing = sum(!is.na(positive_for_drug)),
positive_test = sum(positive_for_drug, na.rm = TRUE),
perc_positive = round(100 * positive_test / n_non_missing, 1)) %>%
select(drug_type, sex, year_cat, perc_positive) %>%
unite(sex_year_cat, sex, year_cat) %>%
spread(sex_year_cat, perc_positive) %>%
knitr::kable(col.names = c("Drug type", "F 1999-2002",
"F 2003-2006", "F 2007-2010",
"M 1999-2002", "M 2003-2006",
"M 2007-2010"))
```

```{r echo=FALSE}
library(ggplot2)
library(ggthemes)
rep_clean_fars<- clean_fars %>%
select(year,agecat,drug_type,positive_for_drug) %>%
dplyr::filter(drug_type!= "Alcohol") %>% 
dplyr::filter(!is.na(agecat)) %>%
 mutate(year=as.factor(year)) %>%
mutate(positive_for_drug=as.logical(positive_for_drug)) %>%
mutate(agecat=as.factor(agecat)) %>%
group_by(year,agecat) %>%
summarize (percent= mean(positive_for_drug)*500) 
```

**Figure 1: Prevalence of nonalcohol drugs in fatally injured drivers  by year and age group**

```{r}
plot1<- ggplot(rep_clean_fars,aes(x = year,y= percent, group = agecat,shape= agecat)) +
geom_point () + 
xlab("Year") + ylab ("Positive for Nonalcohol Drugs %") +
geom_line() +
scale_y_continuous(breaks=c(0,10,20,30)) +
 theme_few() 
 plot1
```

```{r echo=FALSE}
rep_clean_fars_2<-clean_fars %>%
select(year,agecat,drug_type,positive_for_drug) %>%
dplyr::filter(drug_type!= "Alcohol") %>% 
dplyr::filter(!is.na(agecat)) %>%
  mutate(year=as.factor(year)) %>%
  mutate(positive_for_drug=as.logical(positive_for_drug)) %>%
  mutate(drug_type=as.factor(drug_type)) %>%
mutate(agecat=as.factor(agecat)) %>%
group_by(year,drug_type) %>%
summarize(percent= mean(positive_for_drug)*100)
```
**Figure 2: Prevalence of nonalcohol drugs in fatally injured drivers by year and drug type**

```{r}
plot2 <- ggplot(rep_clean_fars_2,aes(x=year,y=percent,group = drug_type,shape=drug_type)) +
geom_point(size = 1.5) + 
xlab("Year") +
ylab ("Positive for Drugs, %") +
scale_y_continuous(breaks=c(0,5,10)) +
geom_line() +
theme_few() 
plot2 
```

```{r echo=FALSE}
rep_clean_fars_3<-clean_fars %>%
select(year,agecat,drug_type,positive_for_drug) %>%
dplyr::filter(drug_type =="Cannabinoid") %>% 
dplyr::filter(!is.na(agecat)) %>%
  mutate(year=as.factor(year)) %>%
  mutate(positive_for_drug=as.logical(positive_for_drug)) %>%
mutate(agecat=as.factor(agecat)) %>%
group_by(year,agecat) %>%
summarize(percent= mean(positive_for_drug)*100)
```

**Figure3: Prevalence of cannabinoid drugs in fatally injured drivers by year and age group**

```{r}
plot3<- ggplot(rep_clean_fars_3,aes(x=year,y=percent,group = agecat,shape=agecat)) +
geom_point (size = 1.5) + 
xlab("Year") +
ylab ("Positive for Cannabinoid, %") +
geom_line() +
scale_y_continuous(breaks=c(0,5,10,15,20)) +
theme_few() 
plot3
```


**Prevalence of drugs in fatally injured drivers for 1999 and 2010 by drug type**

```{r}
library(tidyr)
clean_fars %>%
filter(year %in%c(1999,2010)) %>%
filter(!is.na(year)) %>%
filter(!is.na(drug_type)) %>%
group_by(drug_type, year) %>%
summarize(n = sum(!is.na(positive_for_drug)),
x =sum(positive_for_drug, na.rm = TRUE),
test_result= per_cis(x,n)) %>%
select(drug_type,  year,test_result) %>%
spread(year, test_result) %>%
knitr::kable(col.names = c("Drug type",1999,2010))
```


**Statistics for testing for trend in prevalence of drugs over study years by drug type using
Cochran-Armitage trend test**

```{r}
drug_list <- c("Alcohol", "Nonalcohol", "Narcotic", "Depressant",
"Stimulant", "Cannabinoid", "Other")
drug_trend_tests_ca <- lapply(drug_list, test_trend_ca)
drug_trend_tests_ca <- dplyr::bind_rows(drug_trend_tests_ca) %>%
dplyr::mutate(drug = drug_list) %>%
dplyr::select(drug, Z, p.value)
drug_trend_tests_ca %>% knitr::kable()
```

**Statistics for testing for trend in prevalence of drugs over study years by drug type using Waldtest of logistic regression coefficient for “year”**

```{r}
 drug_list <- c("Alcohol", "Nonalcohol", "Narcotic", "Depressant",
"Stimulant", "Cannabinoid", "Other")
drug_trend_tests_log_reg <- lapply(drug_list, test_trend_log_reg)
drug_trend_tests_log_reg <- dplyr::bind_rows(drug_trend_tests_log_reg) %>%
dplyr::mutate(drug = drug_list) %>%
dplyr::select(drug, Z, p.value)
drug_trend_tests_log_reg %>% knitr::kable()
 
```



